---
title: "data-loading-and-cleaning"
author: "Hayes Waddell"
output: html_document
---

# Data Loading

## Function to clean individual data sets grouped by pitch type and count.

- filepath: string of path to data set
- count_label: string of the count for that data set

This function removes unwanted columns, groups by pitch type and aggregates kept columns, and mutates whiffs and swings into rate stats.
```{r}
# load necessary library
library(dplyr)

# clean data
clean_count_data <- function(filepath, count_label) {
  read.csv(filepath) %>%
    select(pitch_type, pitches, whiffs, swings, xwoba) %>%
    filter(!is.na(xwoba)) %>%
    group_by(pitch_type) %>%
    summarise(
      pitches = sum(pitches),
      whiffs  = sum(whiffs),
      swings  = sum(swings),
      xwoba   = mean(xwoba)
    ) %>%
    filter(pitches > 50) %>%
    mutate(
      count      = count_label,
      whiff_rate = whiffs / swings,
      swing_rate = swings / pitches,
    ) %>%
    select(-whiffs, -swings)
}
```

## Load and clean individual data sets for each count

```{r}
# load data
df_00 <- clean_count_data("data/raw/25_0-0.csv", "0-0")
df_01 <- clean_count_data("data/raw/25_0-1.csv", "0-1")
df_02 <- clean_count_data("data/raw/25_0-2.csv", "0-2")
df_10 <- clean_count_data("data/raw/25_1-0.csv", "1-0")
df_11 <- clean_count_data("data/raw/25_1-1.csv", "1-1")
df_12 <- clean_count_data("data/raw/25_1-2.csv", "1-2")
df_20 <- clean_count_data("data/raw/25_2-0.csv", "2-0")
df_21 <- clean_count_data("data/raw/25_2-1.csv", "2-1")
df_22 <- clean_count_data("data/raw/25_2-2.csv", "2-1")
df_30 <- clean_count_data("data/raw/25_3-0.csv", "3-0")
df_31 <- clean_count_data("data/raw/25_3-1.csv", "3-1")
df_32 <- clean_count_data("data/raw/25_3-2.csv", "3-2")
```

## Join into one data set

```{r}
# join all count data sets into one long data set
df_combined <- bind_rows(df_00, df_01, df_02, df_10, df_11, df_12, df_20, df_21, df_22, df_30, df_31, df_32)
```

## Save all new data sets to csv in data/clean/

```{r}
datasets <- list(
  "00" = df_00, "01" = df_01, "02" = df_02,
  "10" = df_10, "11" = df_11, "12" = df_12,
  "20" = df_20, "21" = df_21, "22" = df_22,
  "30" = df_30, "31" = df_31, "32" = df_32
)

for (name in names(datasets)) {
  write.csv(datasets[[name]], paste0("data/clean/", name, ".csv"), row.names = FALSE)
}

write.csv(df_combined, "data/clean/combined.csv", row.names = FALSE)
```

